Tools for identifying unexpectely low microfossil count sums

### 2019-05-07 Rmd/count_check_MS.Rmd
changed sign of bird analyses

#set ggplot theme
theme_set(theme_bw(base_size = 12))
options(digits = 1)

@@

A tiny minority of the over `r format(floor(bird_singleton_summary$n_routes[1]/1000) * 1000, big.mark = ",")` bird counts from the North American breeding bird survey [@NAmerican_Breeding_Bird2018]
lack singletons at the species level (Table \@ref(tab:bird-table))).
To explore the effect of taxonomic richness on the prevalence of singletons, I aggregate the birds counts to progressively lower taxonomic resolutions.
As richness declines, the proportion of counts lacking singletons increases (Table \@ref(tab:bird-table)), reaching a moderate proportion at the order level, where most of the counts lacking singletons have fewer than five taxa.
The vast majority of counts without singletons have a greatest common divisor of one, except when taxonomic richness is low.

```{r bird-table}
bird_singleton_summary %>%
  select(`Taxonomic level` = taxonomic_level, `Percent without singletons` = pc_singletons, `Percent GCD > 1` = GCD2plus, `Mean richness` = mean_richness) %>%
  knitr::kable(digits = c(0, 4, 4, 3), caption = "Percent of bird counts without singletons and of greatest common divisor (GCD) greater than one of counts without singletons, at different taxonomic levels.")
```

To test the sensitivity of the prevalence of counts without singletons to the count sum, I re-sample the counts with 400 or more observations to smaller count sums using a multivariate hypergeometric distribution.
At the order level, the proportion of counts without singletons declines steeply until it reaches a minimum at about 200, thereafter is rises slowly (Fig. \@ref(fig:bird-recount)).
The prevalence of counts without singletons is high for small counts because it is possible that only the common taxa are counted; with larger counts, the chance of counting a single individual of a rare taxon increases.
With even larger counts the prevalence of counts without singletons increases due to saturation as singletons become doubletons and few new taxa are found.

>>>>

#set ggplot theme
theme_set(theme_bw(base_size = 12))
options(digits = 2)

@@

The vast majority of the over `r format(floor(bird_singleton_summary$n_routes[1]/1000) * 1000, big.mark = ",")` bird counts from the North American breeding bird survey [@NAmerican_Breeding_Bird2018]
have singletons at the species level (Table \@ref(tab:bird-table))).
To explore the effect of taxonomic richness on the prevalence of singletons, I aggregate the birds counts to progressively lower taxonomic resolutions.
As richness declines, the proportion of counts having singletons declines (Table \@ref(tab:bird-table)), reaching a moderate proportion at the order level, where most of the counts lacking singletons have fewer than five taxa.
The vast majority of counts have a greatest common divisor of one, except when taxonomic richness is low.

```{r bird-table}
options(digits = 4)
bird_singleton_summary %>%
  select(`Taxonomic level` = taxonomic_level, `Percent with singletons` = p_singletons, `Percent GCD = 1` = GCD1, `Median richness` = median_richness) %>%
  knitr::kable(digits = c(0, 6, 6, 3), caption = "Percent of bird counts with singletons and of greatest common divisor (GCD) of one, at different taxonomic levels.")
options(digits = 2)
```

To test the sensitivity of the prevalence of counts without singletons to the count sum, I re-sample the counts with 400 or more observations to smaller count sums using a multivariate hypergeometric distribution.
At the order level, the proportion of counts with singletons increases steeply until it reaches a maximum at about 200, thereafter is declines slowly (Fig. \@ref(fig:bird-recount)).
The prevalence of counts with singletons is low for small counts because it is possible that only the common taxa are counted; with larger counts, the chance of counting a single individual of a rare taxon increases.
With even larger counts the prevalence of counts with singletons decreases due to saturation as singletons become doubletons and few new taxa are found.
